---
title: "LDA K-fold cross-validation part 2"
output: html_notebook
---

# Goals

  * Find the optimal number of topics using a cross-validation approach
  * Select a final LDA model using the aggressive TeX removal sparse matrix
  * Learn a bit about parallel cloud computing

Part 1, [07a-lda-kfold-xval.Rmd](https://github.com/phively/uchicago-thesis/blob/master/07a-lda-kfold-xval.Rmd), is too slow to complete locally so I'm going to parallelize the high-topic-count models on AWS.

# Libraries used

```{r, warning=F, message=F, error=F, comment=F}
library(tidyverse)
library(wranglR)
library(tm)
library(slam)
library(topicmodels)
library(foreach)
library(doParallel)
```

# Load the data

See [06-dataset-comparison.Rmd](https://github.com/phively/uchicago-thesis/blob/master/06-dataset-comparison.Rmd) for details on how the sparse matrix was generated.

```{r, message=F}
# Load sparse matrix with no tex
load(file = "data/sparse_matrix_no_tex.zip")
```

# Setup

## Data

Set the number of groups $K$ and generate indices.

```{r}
# Set K
K <- 5

# Create k groups, as close to equal size as possible
set.seed(93401)
rand_ind <- KFoldXVal(1:sparse$nrow, k = K)
```

Create the functions to be used for cross-validation.

```{r}
# Cross-validation matrix generator function, split_set()
source("R/function-split_set.R")

# Parallel cross-validation function
xval <- function(dtm, index, ntopics, ...) {
  # Generate a list of models in parallel
  models <- foreach(i = 1:K, .packages = c("wranglR", "topicmodels", "slam"),
    .export = "split_set") %dopar% {
    # Data to use
    this_ind <- unlist(index)[unlist(index) %nin% index[[i]]]
    this_data <- split_set(dtm, this_ind)
    # Return a topic model
    LDA(this_data, k = ntopics, method = "VEM", control = list(...))
  }
  # Return the list of models
  return(models)
}
```

## Core initialization

Use 12 (of 16) cores for the process.

```{r}
registerDoParallel(12)
getDoParWorkers()
```

# Topic modeling

## 3 topics

```{r}
gc()
now <- Sys.time()
models3 <- xval(dtm = sparse, index = rand_ind, ntopics = 3, seed = 17481, keep = 1)
(t3 <- Sys.time() - now)
save(models3, file = "results/lda_k_3.zip", compress = "xz", compression_level = 1)
```

## 5 topics

```{r}
gc()
now <- Sys.time()
models5 <- xval(dtm = sparse, index = rand_ind, ntopics = 5, seed = 74749, keep = 1)
(t5 <- Sys.time() - now)
save(models5, file = "results/lda_k_5.zip", compress = "xz", compression_level = 1)
```

## 10 topics

```{r}
gc()
now <- Sys.time()
models10 <- xval(dtm = sparse, index = rand_ind, ntopics = 10, seed = 90672, keep = 1)
(t10 <- Sys.time() - now)
save(models10, file = "results/lda_k_10.zip", compress = "xz", compression_level = 1)
```

## 15 topics

```{r}
gc()
now <- Sys.time()
models15 <- xval(dtm = sparse, index = rand_ind, ntopics = 15, seed = 53177, keep = 1)
(t15 <- Sys.time() - now)
save(models15, file = "results/lda_k_15.zip", compress = "xz", compression_level = 1)
```

## 20 topics

```{r}
gc()
now <- Sys.time()
models20 <- xval(dtm = sparse, index = rand_ind, ntopics = 20, seed = 45281, keep = 1)
(t20 <- Sys.time() - now)
save(models20, file = "results/lda_k_20.zip", compress = "xz", compression_level = 1)
```

## 25 topics

```{r}
gc()
now <- Sys.time()
models25 <- xval(dtm = sparse, index = rand_ind, ntopics = 25, seed = 86244, keep = 1)
(t25 <- Sys.time() - now)
save(models25, file = "results/lda_k_25.zip", compress = "xz", compression_level = 1)
```

## 30 topics

```{r}
gc()
now <- Sys.time()
models30 <- xval(dtm = sparse, index = rand_ind, ntopics = 30, seed = 41531, keep = 1)
(t30 <- Sys.time() - now)
save(models30, file = "results/lda_k_30.zip", compress = "xz", compression_level = 1)
```

## 35 topics

```{r}
gc()
now <- Sys.time()
models35 <- xval(dtm = sparse, index = rand_ind, ntopics = 35, seed = 68513, keep = 1)
(t35 <- Sys.time() - now)
save(models35, file = "results/lda_k_35.zip", compress = "xz", compression_level = 1)
```

## 40 topics

```{r}
gc()
now <- Sys.time()
models40 <- xval(dtm = sparse, index = rand_ind, ntopics = 40, seed = 19275, keep = 1)
(t40 <- Sys.time() - now)
save(models40, file = "results/lda_k_40.zip", compress = "xz", compression_level = 1)
```

## 45 topics

```{r}
gc()
now <- Sys.time()
models45 <- xval(dtm = sparse, index = rand_ind, ntopics = 45, seed = 68513, keep = 1)
(t45 <- Sys.time() - now)
save(models45, file = "results/lda_k_45.zip", compress = "xz", compression_level = 1)
```

# Perplexities

## In-sample results

```{r}
# Set up data frame and labels
perplexity <- data.frame(
  model = c(3, 5, 10, 15, 20, 25, 30, 35, 40, 45) %>% rep(each = K),
  fold = rep(1:K, times = 11),
  # Add time elapsed in hours
  time = c(t3, t5, t10, t15, t20, t25, t30, t35, t40, t45) %>% as.numeric(units = "hours") %>% rep(each = K)
)

# Calculate in-sample perplexity
in_sample = c(
  lapply(models3, FUN = perplexity) %>% unlist(),
  lapply(models5, FUN = perplexity) %>% unlist(),
  lapply(models10, FUN = perplexity) %>% unlist(),
  lapply(models15, FUN = perplexity) %>% unlist(),
  lapply(models20, FUN = perplexity) %>% unlist(),
  lapply(models25, FUN = perplexity) %>% unlist(),
  lapply(models30, FUN = perplexity) %>% unlist(),
  lapply(models35, FUN = perplexity) %>% unlist(),
  lapply(models40, FUN = perplexity) %>% unlist(),
  lapply(models45, FUN = perplexity) %>% unlist()
)
perplexity <- perplexity %>% mutate(in_sample)
```

## Out-of-sample results

```{r}
# Function for out-of-sample perplexity
oos_perplexity <- function(models_list) {
  foreach (i = 1:K) %dopar% {
    perplexity(models_list[[i]], newdata = split_set(sparse, rand_ind[[i]]))
  }
}

# Calculate out-of-sample perplexity
perp3 <- oos_perplexity(models3) %>% unlist()
perp5 <- oos_perplexity(models5) %>% unlist()
perp10 <- oos_perplexity(models10) %>% unlist()
perp15 <- oos_perplexity(models15) %>% unlist()
perp20 <- oos_perplexity(models20) %>% unlist()
perp25 <- oos_perplexity(models25) %>% unlist()
perp30 <- oos_perplexity(models30) %>% unlist()
perp35 <- oos_perplexity(models35) %>% unlist()
perp40 <- oos_perplexity(models40) %>% unlist()
perp45 <- oos_perplexity(models45) %>% unlist()

# Append to perplexity data frame
perplexity <- perplexity %>% mutate(
  out_sample = c(perp3, perp5, perp10, perp15, perp20, perp25, perp30, perp35, perp40) %>% unlist()
)

# Save results
write.csv(perplexity, file = "results/lda_k_perp.csv", row.names = FALSE)
```

# Additional topics around the observed minimum

## 38 topics

```{r}
gc()
now <- Sys.time()
models38 <- xval(dtm = sparse, index = rand_ind, ntopics = 38, seed = 14508, keep = 1)
(t38 <- Sys.time() - now)
save(models38, file = "results/lda_k_38.zip", compress = "xz", compression_level = 1)
```

## 39 topics

```{r}
gc()
now <- Sys.time()
models39 <- xval(dtm = sparse, index = rand_ind, ntopics = 39, seed = 85390, keep = 1)
(t39 <- Sys.time() - now)
save(models39, file = "results/lda_k_39.zip", compress = "xz", compression_level = 1)
```

## 41 topics

```{r}
gc()
now <- Sys.time()
models41 <- xval(dtm = sparse, index = rand_ind, ntopics = 41, seed = 87837, keep = 1)
(t41 <- Sys.time() - now)
save(models41, file = "results/lda_k_41.zip", compress = "xz", compression_level = 1)
```

## 42 topics

```{r}
gc()
now <- Sys.time()
models42 <- xval(dtm = sparse, index = rand_ind, ntopics = 42, seed = 81348, keep = 1)
(t42 <- Sys.time() - now)
save(models42, file = "results/lda_k_42.zip", compress = "xz", compression_level = 1)
```

## 50 topics

```{r}
gc()
now <- Sys.time()
models50 <- xval(dtm = sparse, index = rand_ind, ntopics = 50, seed = 68513, keep = 1)
(t50 <- Sys.time() - now)
save(models50, file = "results/lda_k_50.zip", compress = "xz", compression_level = 1)
```

# New perplexity

```{r}
perps_new <- data.frame(
  model = c(38, 39, 41, 42, 50) %>% rep(each = K),
  fold = rep(1:K, times = 5),
  # Add time elapsed in hours
  time = c(t38, t39, t41, t42, t50) %>% as.numeric(units = "hours") %>% rep(each = K)
)

in_sample = c(
  lapply(models38, FUN = perplexity) %>% unlist(),
  lapply(models39, FUN = perplexity) %>% unlist(),
  lapply(models41, FUN = perplexity) %>% unlist(),
  lapply(models42, FUN = perplexity) %>% unlist(),
  lapply(models50, FUN = perplexity) %>% unlist()
)
perps_new <- perps_new %>% mutate(in_sample)
```


## Out-of-sample

```{r}
perp38 <- oos_perplexity(models38) %>% unlist()
perp39 <- oos_perplexity(models39) %>% unlist()
perp41 <- oos_perplexity(models41) %>% unlist()
perp42 <- oos_perplexity(models42) %>% unlist()
perp50 <- oos_perplexity(models50) %>% unlist()

perps_new <- perps_new %>% mutate(out_sample = c(perp38, perp39, perp41, perp42, perp50))

perplexity <- rbind(perplexity, perps_new) %>% arrange(model, fold)

write.csv(perplexity, file = "results/lda_k_perp.csv")
```

# Perplexity comparison

```{r}
perplexity %>% filter(model >= 10) %>% group_by(model) %>% mutate(in_avg = mean(in_sample), out_avg =mean(out_sample)) %>%
  ggplot(aes(x = model)) + geom_point(aes(y = in_sample, color = "In-sample")) + geom_line(aes(y = in_avg, color = "In-sample")) +
  geom_point(aes(y = out_sample, color = "Out-sample")) + geom_line(aes(y = out_avg, color = "Out-sample")) +
  labs(x = "topics", y = "perplexity", color = "", title = paste(K, "-fold cross-validation perplexity", sep = ""))
```

The minimum is at $K = 40$, which is conveniently a nice round number.

# Final model

```{r}
gc()
now <- Sys.time()
models40f <- LDA(sparse, k = 40, method = "VEM",
                 control = list(seed = 19275, keep = 1, verbose = 1))
(t40f <- Sys.time() - now)
save(models40f, file = "results/lda_final_40.zip", compress = "xz", compression_level = 1)
```