---
title: "Correlated topic models K"
output: html_notebook
---

# Goals

  * Identify a suitable number of topics $K$ to use for the CTM.

# Libraries used

```{r, warning=F, message=F, error=F, comment=F}
library(tidyverse)
library(slam)
library(stm)
library(foreach)
library(doParallel)
```

Set up parallel runs

```{r}
registerDoParallel(5)
getDoParWorkers()
```

# Load the final data

See [06-dataset-comparison.Rmd](https://github.com/phively/uchicago-thesis/blob/master/06-dataset-comparison.Rmd) for details on how the sparse matrix was generated.

```{r, message=F}
# Load sparse matrix with no tex
load(file = "data/sparse_matrix_no_tex.zip")
sparse <- readCorpus(sparse, type = "slam")
```

# CTM method without correlations

We can get a model like an LDA by enforcing a diagonal $\Sigma$ with `sigma.prior = 1` but note that the topic distribution in this model is not actually a Dirichlet.

```{r}
nc_ctm40 <- stm(sparse$documents, sparse$vocab, K = 40, init.type = "Spectral", ngroups = 7,
  control = list(maxV = length(sparse$vocab)), sigma.prior = 1)
nc_ctm40$time/60
```

# Rerun the 40 topic model as CTM

Correlated topic models relax the $\Sigma$ is diagonal constraint.

```{r}
now <- Sys.time()
ctm40 <- stm(sparse$documents, sparse$vocab, K = 40, init.type = "Spectral", ngroups = 7,
  control = list(maxV = length(sparse$vocab)))
(t40 <- Sys.time() - now)
```

# Compare diagonal and non-diagonal $\Sigma$

```{r, fig.height=8}
# Time to complete
nc_ctm40$time/60
ctm40$time/60
# Correlations
plot(topicCorr(nc_ctm40))
plot(topicCorr(ctm40))
# Most frequent topics
plot(nc_ctm40, n = 5, xlim = c(0, .1))
plot(ctm40, n = 5, xlim = c(0, .1))
```

# Other numbers of topics

That was surprisingly fast. Try every 5 as before.

```{r}
ctm5  <- stm(sparse$documents, sparse$vocab, K = 5,  init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm10 <- stm(sparse$documents, sparse$vocab, K = 10, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm15 <- stm(sparse$documents, sparse$vocab, K = 15, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm20 <- stm(sparse$documents, sparse$vocab, K = 20, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm25 <- stm(sparse$documents, sparse$vocab, K = 25, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm30 <- stm(sparse$documents, sparse$vocab, K = 30, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm35 <- stm(sparse$documents, sparse$vocab, K = 35, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm45 <- stm(sparse$documents, sparse$vocab, K = 45, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
ctm50 <- stm(sparse$documents, sparse$vocab, K = 50, init.type = "Spectral", ngroups = 7, control = list(maxV = length(sparse$vocab)))
```

# Multiple K

```{r}
now <- Sys.time()
manyK <- foreach(this.k = seq(50, 5, by = -5), .packages = "stm") %dopar% {
  output <- searchK(sparse$documents, sparse$vocab, K = this.k, init.type = "Spectral", ngroups = 7,
  proportion = .5, heldout.seed = 13316, verbose = FALSE)
  return(output$results)
}
(tK <- Sys.time() - now)
```

## searchK results

```{r, fig.height=8}
manyK <- list(results = bind_rows(manyK), call = NULL)
attr(manyK, "class") <- "searchK"
plot(manyK)
```

## Try additional K

```{r}
now <- Sys.time()
moreK <- foreach(this.k = seq(100, 55, by = -5), .packages = "stm") %dopar% {
  output <- searchK(sparse$documents, sparse$vocab, K = this.k, init.type = "Spectral", ngroups = 7,
  proportion = .5, heldout.seed = 13316, verbose = FALSE)
  return(output$results)
}
(tK2 <- Sys.time() - now)
```

```{r, fig.height=8}
tmp <- manyK
tmp$results <- rbind(bind_rows(moreK), manyK$results)
plot(tmp)
```

# Save results

```{r}
write.csv(rbind(manyK$results, bind_rows(moreK)), file = "results/ctm_holdout.csv")
```

```{r}
save(ctm5, ctm10, ctm15, ctm20, ctm25, ctm30, ctm35, ctm40, ctm45, ctm50, nc_ctm40, manyK, moreK,
     file ="results/ctm_by_5.zip", compress = "xz", compression_level = 1)
```