---
title: "Multicore test"
output: html_document
---

# Setup

```{r}
# Libraries
library(foreach)
library(doParallel)
library(wranglR)
library(dplyr)
library(topicmodels)
library(slam)

# Backend
registerDoParallel(3)
```

# Data

```{r}
# Iris data
x <- iris[which(iris[,5] != "setosa"), c(1,5)]
trials <- 5000

# Test DTM data
source("R/function-split_set.R")
load("data/sparse_test.zip")
K <- 3
set.seed(123)
inds <- KFoldXVal(1:sparse_test$nrow, k = K)
```

# Iris

## No thread

t2.micro    = 60.465 seconds
Sandybridge = 28.460 seconds
c4.2xlarge  = 10.945 seconds

```{r}
print("Iris %do%")
ptime <- system.time({
  r <- foreach(icount(trials), .combine=cbind) %do% {
    ind <- sample(100, 100, replace=TRUE)
    result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
    coefficients(result1)
  }
})[3]
print(ptime)
```

## Thread

t2.micro    = 60.518 seconds (not threaded)
Sandybridge = 18.920 seconds
c4.2xlarge  =  4.102 seconds

```{r}
print("Iris %dopar%")
ptime <- system.time({
  r <- foreach(icount(trials), .combine=cbind) %dopar% {
    ind <- sample(100, 100, replace=TRUE)
    result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
    coefficients(result1)
  }
})[3]
print(ptime)
```

# LDA

## No thread

Sandybridge = 127.890 seconds
c4.2xlarge  =  87.163 seconds

xval function, no threading

```{r}
xval <- function(dtm, index, ntopics, seed, ...) {
  # Generate a list of models
  models <- foreach(i = 1:K, .packages = c("wranglR", "topicmodels", "slam"),
    .export = "split_set") %do% {
    # Data to use
    this_ind <- unlist(index)[unlist(index) %nin% index[[i]]]
    this_data <- split_set(dtm, this_ind)
    # Return a topic model
    return(
      LDA(this_data, k = ntopics, method = "VEM",
          control = list(seed = seed, ...))
    )
  }
  # Return the list of models
  return(models)
}
```
```{r}
print("LDA %do%")
now <- Sys.time()
r <- xval(sparse_test, inds, ntopics = 5, seed = 123)
print(Sys.time() - now)
```

## Thread

Sandybridge = 52.100 seconds
c4.2xlarge  = 32.503 seconds

xval function, threading

```{r}
xval <- function(dtm, index, ntopics, ...) {
  # Generate a list of models in parallel
  models <- foreach(i = 1:K, .packages = c("wranglR", "topicmodels", "slam"),
    .export = "split_set") %dopar% {
    # Data to use
    this_ind <- unlist(index)[unlist(index) %nin% index[[i]]]
    this_data <- split_set(dtm, this_ind)
    # Return a topic model
    LDA(this_data, k = ntopics, method = "VEM",
        control = list(...))
  }
  # Return the list of models
  return(models)
}
```
```{r}
print("LDA %dopar%")
now <- Sys.time()
r <- xval(sparse_test, inds, ntopics = 5, seed = 123)
print(Sys.time() - now)
```

# Many-core test

Sandybridge = 16.830 seconds
c4.2xlarge  =  2.719 seconds

```{r}
# Allow 8 workers
registerDoParallel(8)

# Bootstrapping
print("Iris many %dopar%")
ptime <- system.time({
  r <- foreach(icount(trials), .combine=cbind) %dopar% {
    ind <- sample(100, 100, replace=TRUE)
    result1 <- glm(x[ind,2]~x[ind,1], family=binomial(logit))
    coefficients(result1)
  }
})[3]
print(ptime)
```

# Real-world test

Here's a real-world comparison of the full dataset, 5-fold cross-validation of 3 topics.

Sandybridge = 406.202 seconds (single-threaded)
c4.2xlarge  = 177.735 seconds (five threads)

```{r}
# New dataset
load(file = "data/sparse_matrix_no_tex.zip")

# Setup
K <- 5
set.seed(93401)
rand_ind <- KFoldXVal(1:sparse$nrow, k = K)

registerDoSEQ() # Sandybridge
registerDoParallel(K) # c4.2xlarge

# Benchmark
now <- Sys.time()
models3 <- xval(dtm = sparse, index = rand_ind, ntopics = 3, seed = 17481, keep = 1)
(t3 <- Sys.time() - now)
save(models3, file = "results/lda_k_3.zip")
```

Well, that ought to help! That's over 125% speedup. Back to our regularly scheduled cross-validation.